---
title: "Reproducable data Peer Project 1"
author: "garyf"
date: "Saturday, February 14, 2015"
output: html_document
---

This is the peer assessed project one. I have a long history with relational databases. (My favorite SAS PROC is Proc SQL) I am using the R equivilent R library. sqldf

I also discovered the iterator library, I think this will be most helpful ...

I am doing both the SQL way and the R way. I am seeing some differences and I am not sure if the SQL way is the best way to query a data.frame. Please watch for this as there might be 2 analysis per question / part...

# Initial analysis, daily histogram.

```{r}
require(sqldf)
require(iterators)

fitbit <- read.csv("data1/activity.csv")
groupSum <- sqldf("select date, sum(steps) as sum_steps from fitbit where steps != 'NA' group by date")

minAverage <- sqldf("select interval, avg(steps) as avg_steps from fitbit where steps != 'NA' group by interval")
 colors = c( "yellow", "green", "violet", "orange",  "blue", "pink", "cyan") 

 hist( groupSum$sum_steps ,    # apply the hist function 
  breaks=53,  # number of observations over 2 months
   right=FALSE,    # intervals closed on the left 
   col=colors,     # set the color palette 
   main="Steps per Day", # the main title 
   xlab="Day")       # x-axis label 

```

The number of steps by time of day; grouped by 5 minute groups

```{r }
 plot(x=minAverage$interval, y=minAverage$avg_steps)
```

# The R way

 Get the list of unique dates, subset the days then calculate the mean and median, make a new dataframe, print that one at the end.

### If I could loop over the dates without error ... see tryCatch code most useful
### errors in the code prevent knitr from knitting ... this allows the process to continue
### Granted it would be better to halt the while loop when the iterator hasMore == FALSE

 

```{r} 
# Subset the dates
myUniqueDates = unique(fitbit$date)

# empty new data.frame
myMeanDataFrame<-NULL;

# make the iterator, loop forever or until it dies (sad)
myDateIter <-iter(myUniqueDates)
while(TRUE){
 
  d=try(nextElem(myDateIter))
  if (class(d) == "try-error") break  
  
  m1 <- mean(subset(fitbit, fitbit$date==d)$steps) 
  m2 <- median(subset(fitbit, fitbit$date==d)$steps)
  rbind(myMeanDataFrame, data.frame(date=d, mean=m1, median=m2) ) -> myMeanDataFrame
                
}
print("# Fit bit Means and Medians")
print(myMeanDataFrame)
```

## This was based on the SQL commands and appears to be wrong .. me or SQL?
The maximum average steps occur at the timepoint 8:35, with 206 steps by visual inspection. Computed result follows.


```{r }
maxTimePoint <- sqldf('select interval, max(avg_steps) as mostSteps from minAverage')
'Max average STeps'
maxTimePoint$mostSteps
'occured at'
maxTimePoint$interval

```
# Begin the Imputation
## Primary Strategy: Use the per 5 minute block mean to replace the NA's for the same block. Code follows.

Getting the number of NA's can be performed with SQL.
hmmm does not seem to be working ...
```{r}
countNA <- sqldf("select count(steps) from fitbit where steps = 'NA'")
  
countNA
```

## I would like an iterator, found this
### itertools and iterators
[http://www.r-bloggers.com/iterators-in-r/]( Iterators at http://www.r-bloggers.com/iterators-in-r/ )
## This is horrible, I don't like it but it seems to work ...
## There is probably a more R way to do this; ran out of time
### My iterator worked, I could step over the date but I had trouble
### catching the end of the data error, killing the knitr process.

```{r} 

#myCountNA <- function(myDataFrameVector){ 
#  tryCatch(
#    cntr = 0
#    stepsIter <- iter(myDataFrameVector, by="row") 
#      while(1==1) {
#       nxtRow <- nextElem(stepsIter) 
#       if ( is.na(nxtRow$steps) ) { 
#       cntr = cntr + 1 
#      }
#      return(cntr)
#    },
#    error = function(e){return(cntr)}
#  )
#} # Ends the function definitions

#myCountNA(fitbit)
sprintf("There are %d NA's in the original data.",2304)

print("My deepest apologies, I think I ran out of time....")
```