---
title: "Reproducable data Peer Project 1"
author: "garyf"
date: "Saturday, February 14, 2015"
output: html_document
---

This is the peer assessed project one. I have a long history with relational databases. (My favorite SAS PROC is Proc SQL) I am using the R equivilent R library. sqldf

Initial analysis, daily histogram.

```{r}
require(sqldf)
require(iterators)

fitbit <- read.csv("data1/activity.csv")
groupSum <- sqldf("select date, sum(steps) as sum_steps from fitbit where steps != 'NA' group by date")

minAverage <- sqldf("select interval, avg(steps) as avg_steps from fitbit where steps != 'NA' group by interval")
 colors = c( "yellow", "green", "violet", "orange",  "blue", "pink", "cyan") 

 hist( groupSum$sum_steps ,    # apply the hist function 
  breaks=53,  # number of observations over 2 months
   right=FALSE,    # intervals closed on the left 
   col=colors,     # set the color palette 
   main="Steps per Day", # the main title 
   xlab="Day")       # x-axis label 

```

The number of steps by time of day; grouped by 5 minute groups

```{r }
 plot(x=minAverage$interval, y=minAverage$avg_steps)
```

 The maximum average steps occur at the timepoint 8:35, with 206 steps by visual inspection. Computed result follows.
```{r }
maxTimePoint <- sqldf('select interval, max(avg_steps) as mostSteps from minAverage')
'Max average STeps'
maxTimePoint$mostSteps
'occured at'
maxTimePoint$interval

```
Getting the number of NA's can be performed with SQL.
hmmm does not seem to be working ...
```{r}
countNA <- sqldf("select count(steps) from fitbit where steps = 'NA'")
  
countNA
```

## I would like an iterator, found this
### itertools and iterators
[http://www.r-bloggers.com/iterators-in-r/]( Iterators at http://www.r-bloggers.com/iterators-in-r/ )
## This is horrible, I don't like it but it seems to work ...
## There is probably a more R way to do this; ran out of time
```{r} 
myCountNA <- function(myDataFrameVector){ 
  tryCatch(
    cntr = 0
    stepsIter <- iter(myDataFrameVector, by="row") 
      while(1==1) {
       nxtRow <- nextElem(stepsIter) 
       if ( is.na(nxtRow$steps) ) { 
       cntr = cntr + 1 
      }
      return(cntr)
    },
    error = function(e){return(cntr)}
  )
} # Ends the function definitions

myCountNA(fitbit)
```